---
title: 'Performance of GT8Seq amplicon MiSeq data'
author: "Chloe"
output:
  html_document:
    toc: yes
    number_sections: no
    toc_float: yes
    code_folding: show
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE)
```

# Content

1.  Importing and handling tables in CIGAR format in R environment.
2.  Preliminary sequencing yield by Run and Variable1.
3.  Identification of potential genotyping errors.
4.  Read depth yield per sample and locus.
5.  Retention of loci and samples for subsequent analysis.

This tutorial is based on the output of the Dada2-based denoising pipeline. The main output (the cigar table) of the denoising pipeline has the following structure:

$\begin{array}{c|c:c:c:c:c:c} \text{sampleID}&Gene_1,Allele_1&Gene_1,Allele_2&...&Gene_1,Allele_k&... & Gene_m, Allele_{k_m}\\ \hline ID_1 & \text{Read counts} &&& \\ \hdashline ... &&&&\\ \hdashline ID_n &&&& \end{array}$

Where the $Allele$ is coded in Pseudo-CIGAR format, typing "." for the
reference allele, $[0-9]*[A,T,C,G]$ for each point mutation observed,
and $[0-9]*[D,I]=[A,T,C,G]*$ for indels (Insertions and Deletions). The
numbers before the letters denotes the position in the amplicon where
the polymorphism is located.

Documentation of the denoising pipeline can be found in the repository  [malaria-amplicon-pipeline](https://github.com/broadinstitute/malaria-amplicon-pipeline), and the steps used to generate the inputs files for the current tutorial are in [Intro to MHap-Analysis pipeline](https://paulonvnv.github.io/MHap-Analysis/MHap_Analysis_pipeline.html)

## Importing and handling tables in CIGAR format in R environment

Our first step will be to call all required packages and functions in
the R environment:

```{r 'Loading required packages and functions'}
source('/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/functions_and_libraries/amplseq_required_libraries.R')
source('/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/functions_and_libraries/amplseq_functions.R')
```

In the previous section the Pseudo-CIGAR tables were stored in the
computer using the following path structure:

`"STUDYNAME/RUNNAME/dada2/run_dada2/CIGARVariants.out.tsv"`

As this structure is maintained across different sequencing runs, we
have created a function called `read_cigar_tables` that at least only
requires two arguments: `paths` (name of the folder of the study or the
`STUDYNAME`), and `sample_id_pattern` (a pattern string in the IDs of
the samples that differentiate them from controls). In case you don't
want to differentiate samples from controls, set this argument equals to
`"."`. For this particular example, the path will be
`"data/sequencing_data/"` , while the pattern will be `"."`. In case all
`".tvs"` files are stored in one folder, you can use the argument
`cigar_files = list_of_files` instead of the argument `paths`, where
`list_of_files` is a vector with the names of all files we want to read.

```{r 'Uploading data sets'}
cigar_object = read_cigar_tables(paths = "/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq", sample_id_pattern = ".")

##correct well mix-ups
cigar_object@metadata %<>% mutate(Sample_id2 = case_when(
  Sample_id == "G-60p-8plex-1" ~ "G-60p-No-RIPR-VSP11-3",
  Sample_id == "G-60p-8plex-2" ~ "G-60p-No-RIPR-VSP11-2",
  Sample_id == "G-60p-8plex-3" ~ "G-60p-No-RIPR-VSP11-1",
  Sample_id == "G-60p-No-DBP-1-RBP2b-3-1" ~ "G-60p-No-RBP2b-3-FRM1-1-3",
  Sample_id == "G-60p-No-DBP-1-RBP2b-3-2" ~ "G-60p-No-RBP2b-3-FRM1-1-2",
  Sample_id == "G-60p-No-DBP-1-RBP2b-3-3" ~ "G-60p-No-RBP2b-3-FRM1-1-1",
  Sample_id == "G-60p-No-RBP2b-3-FRM1-1-1" ~ "G-60p-No-DBP-1-RBP2b-3-3",
  Sample_id == "G-60p-No-RBP2b-3-FRM1-1-2" ~ "G-60p-No-DBP-1-RBP2b-3-2",
  Sample_id == "G-60p-No-RBP2b-3-FRM1-1-3" ~ "G-60p-No-DBP-1-RBP2b-3-1",
  Sample_id == "G-60p-No-RIPR-VSP11-1" ~ "G-60p-8plex-3",
  Sample_id == "G-60p-No-RIPR-VSP11-2" ~ "G-60p-8plex-2",
  Sample_id == "G-60p-No-RIPR-VSP11-3" ~ "G-60p-8plex-1",
  Sample_id == "HRP-0-05-8plex-3" ~ "HRP-0-05p-8plex-3",
  .default = Sample_id
), Sample_id = Sample_id2)

cigar_object@metadata$Sample_id2 = NULL

rownames(cigar_object@metadata) = cigar_object@metadata$Sample_id
rownames(cigar_object@cigar_table) = cigar_object@metadata$Sample_id

```

Use the function `slotNames` to visualize the slots that the
`cigar_object` has:

```{r}
slotNames(cigar_object)
```

This function generates a S4 cigar object containing three data.frames
and one list:

1.  `cigar_table`, where genotyping information is stored in cigar
    format;

2.  `metadata` table with 4 variables (columns), **a)** `Sample_id`
    which contains samples Id's; **b)** `run` which contains the run
    number (or the name of the folter where the fastq's for that run
    wehere stored), information that is useful for comparing performance
    across batches of samples; **c)** `order_in_plate`, useful to
    identify rare amplifications patterns in the plate; and **d)**
    `typeofSamp` which differentiates between controls and samples of
    interest.

3.  `asv_table` that contains summary metrics of each `asv` and the
    corresponding `cigar string`.

4.  `asv_seqs` which is the nucleotide sequence of each `asv`.

To view these slots use `View(cigar_table@cigar_table)` in the console.

**NOTE:** If there are duplicated samples, only the replicated with the
highest read depth across all loci will be kept.

## Adding Metadata

Metadata would come from different sources, sometimes the sample ID
incorporates metadata in their coding system, other times metadata can
be stored in a different table, having one column specifying the sample
ID.

Inspect the metadata slot and check if the IDs tell you something about
the procidence of the samples

```{r, eval = FALSE}
View(cigar_object@metadata)
```

```{r}
cigar_object@asv_seqs
cigar_object@cigar_table
cigar_object@metadata
cigar_object@asv_table
```


Samples that starts with SP comes from Colombia, samples that starts
with G are fom Guyana, samples that starts with PF or contains GZ are
from Honduras, and samples that starts with HRP are from Venezuela.

We can use the function `mutate` and a regular expression pattern in
`grepl` to adding the country from where the sample was collected to the
metadata. Also update the column typeofSamp, to be able to distinguish
between samples of interest and controls

```{r 'Adding Country of collection'}

cigar_object@metadata %<>% mutate(
  typeofSamp = case_when(
    grepl('-Negative', Sample_id) ~ 'NoTargetControl',
    grepl('-HumanNegative', Sample_id) ~ 'NegativeControl',
    .default = 'SampofInterest'
  ),
  BiologicalRep = case_when(
    grepl('^G', Sample_id) ~ 'G4G2463',
    grepl('^HRP', Sample_id) ~ 'HRP0378'
  ),
  Multiplex.cat = case_when(
      grepl('-8plex', Sample_id) ~ '8plex',
      grepl('-No-DBP-1-RBP2b-3', Sample_id) ~ '6plex1',
      grepl('-No-RBP2b-3-FRM1-1', Sample_id) ~ '6plex2',
      grepl('-No-RIPR-VSP11', Sample_id) ~ '6plex3',
      grepl('-PIGM-RIPR-VSP11-CG2-related', Sample_id) ~ '4plex1',
      grepl('-PIGM-RIPR-DBP-1-RBP2b-3', Sample_id) ~ '4plex2'
  )
)

View(cigar_object@metadata)
```

Use View again to visualize those changes.

The cigar_table is in wide format, and the alleles corresponding to the
sample target are not in consecutive order. Because of that e are going
to convert our cigar_table to a more manageable (easy to read for the
user) format called ampseq format, where amplicons or genes are going to
be in columns, samples in rows, and the allele in cigar format is going
to be in each cell of the matrix together with the read count of that
allele in the sample.

$\begin{array}{c|c:c:c:c} \text{sampleID}&Gene_1&Gene_2&...& Gene_m\\ \hline ID_1 & Allele_{i_{i \in Alleles_{Gene_1}}}\text{:Read count} &&& \\ \hdashline ... &&&&\\ \hdashline ID_n &&&& \end{array}$

For that purpose we are going to use the function `cigar2ampseq`. This
function requires 4 arguments. First a `cigar_object` (containing the
cigar_table and the metadata). Then `markers`, which is a table with the
set of markers and their names (in a column named `amplicon`). This
argument is optional, and in case is `NULL`, the name of the amplicons
will be extracted from the columns in the `cigar_object`, but
information regarding chormosome location (information required to
estimate relatedness under the hmmIBD algorithm) and length of the
amplicon will be missing. The next arguments are `min_abd` and
`min_ratio`, which are the minimum abundance (minimum read count)
required to call an allele and the minimum ratio between the minor and
major alleles in a polyclonal sample. By default these two arguments are
10 and 0.1 respectively. The last argument is `remove_controls`, by
default is `FALSE`, but if change to `TRUE` the information of the
controls will be stored in a separated slot. Thus, this function
automatically differentiate between controls and samples of interest,
and its output is a `ampseq_object` S4 list with 11 slots: `gt` where
the genotypic information of the samples of interest is stored in a
`ampseq` format; the `asv_table` with summary metrics and the cigar
strings for each asv (allele); the `asv_seqs` list that store the
nucleotide sequence of each asv, the `metadata` of the samples of
interest; `controls` with all the information of the controls; `markers`
containing the information of the markers such as the chromosome
location; and other empty slots required for the next steps such as
`loci_performance` and `pop_summary`.

For the next steps we are going to convert the cigar_object into an
ampseq_object in two different ways:

1.  Keeping controls and all asvs observed in the data set.
2.  Removing controls and asvs with less than 5 reads of support.

But first lest upload the information of the markers into the R
environment:

```{r}
markers = read.csv('/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/reference/Pviv_P01/PvGTSeq8_markersTable.csv')
```

Inspect the table using View() in the console.

```{r}
ampseq_object_abd1 = cigar2ampseq(cigar_object, markers = markers, min_abd = 1, min_ratio = 0.001, remove_controls = F)
ampseq_object = cigar2ampseq(cigar_object, markers = markers, min_abd = 10, min_ratio = 0.000001, remove_controls = F) ##default was min_abd = 10, min_ratio 0.1
```


## Preliminary sequencing yield by Run and Country

Sample success is defined as the percentage of samples which amplify a
specific percentage of loci given the allele detection coverage
threshold or number of paired reads that support the allele or ASV in a
sample. Here we are going to show 6 different coverage thresholds: 1, 5,
10, 20, 50, and 100 paired reads.

### Overall sequencing yield

Measure the read coverage by using the function get_ReadDepth_coverage:

```{r}
ReadDepth_coverage = get_ReadDepth_coverage(ampseq_object_abd1, variable = NULL)
```

For each sample, calculates how many loci has a read depth equals or
greater than 1, 5, 10, 20, 50, or 100:

```{r}
sample_performance = ReadDepth_coverage$plot_read_depth_heatmap$data %>%
    mutate(Read_depth = case_when(
      is.na(Read_depth) ~ 0,
      !is.na(Read_depth) ~ Read_depth
    )) %>%
    summarise(amplified_amplicons1 = sum(Read_depth >= 1)/nrow(ampseq_object_abd1@markers),
              amplified_amplicons5 = sum(Read_depth >= 5)/nrow(ampseq_object_abd1@markers),
              amplified_amplicons10 = sum(Read_depth >= 10)/nrow(ampseq_object_abd1@markers),
              amplified_amplicons20 = sum(Read_depth >= 20)/nrow(ampseq_object_abd1@markers),
              amplified_amplicons50 = sum(Read_depth >= 50)/nrow(ampseq_object_abd1@markers),
              amplified_amplicons100 = sum(Read_depth >= 100)/nrow(ampseq_object_abd1@markers),
              .by = Sample_id) %>%
    pivot_longer(cols = starts_with('amplified_amplicons'), values_to = 'AmpRate', names_to = 'Threshold') %>%
    mutate(Threshold = as.integer(gsub('amplified_amplicons', '', Threshold)))


```

Generate a plot showing the percentage of loci (amplicon or marker)
which contains alleles that pass the thresholds in the x-axis, and the
percentage of samples which amplify a specific percentage of loci given
the allele detection thresholds in the y-axis. Include a vertical line
that indicates samples that has amplify at least 50% of the loci with
the especific read depth coverage.

```{r}
plot_precentage_of_samples_over_min_abd = sample_performance %>%
    summarise(AmpRate5 = round(100*sum(AmpRate >= .05)/n(), 1),
              AmpRate10 = round(100*sum(AmpRate >= .10)/n(), 1),
              AmpRate15 = round(100*sum(AmpRate >= .15)/n(), 1),
              AmpRate20 = round(100*sum(AmpRate >= .20)/n(), 1),
              AmpRate25 = round(100*sum(AmpRate >= .25)/n(), 1),
              AmpRate30 = round(100*sum(AmpRate >= .30)/n(), 1),
              AmpRate35 = round(100*sum(AmpRate >= .35)/n(), 1),
              AmpRate40 = round(100*sum(AmpRate >= .40)/n(), 1),
              AmpRate45 = round(100*sum(AmpRate >= .45)/n(), 1),
              AmpRate50 = round(100*sum(AmpRate >= .50)/n(), 1),
              AmpRate55 = round(100*sum(AmpRate >= .55)/n(), 1),
              AmpRate60 = round(100*sum(AmpRate >= .60)/n(), 1),
              AmpRate65 = round(100*sum(AmpRate >= .65)/n(), 1),
              AmpRate70 = round(100*sum(AmpRate >= .70)/n(), 1),
              AmpRate75 = round(100*sum(AmpRate >= .75)/n(), 1),
              AmpRate80 = round(100*sum(AmpRate >= .80)/n(), 1),
              AmpRate85 = round(100*sum(AmpRate >= .85)/n(), 1),
              AmpRate90 = round(100*sum(AmpRate >= .90)/n(), 1),
              AmpRate95 = round(100*sum(AmpRate >= .95)/n(), 1),
              AmpRate100 = round(100*sum(AmpRate >= 1)/n(), 1),
              .by = c(Threshold)
    ) %>%
    pivot_longer(cols = paste0('AmpRate', seq(5, 100, 5)),
                 values_to = 'Percentage',
                 names_to = 'AmpRate') %>%
    mutate(AmpRate = as.numeric(gsub('AmpRate','', AmpRate)))%>%
    ggplot(aes(x = AmpRate, y = Percentage, color = as.factor(Threshold), group = as.factor(Threshold))) +
    geom_line() +
    geom_vline(xintercept = 50, linetype = 2) +
    theme_bw() +
    labs(x = '% of amplified loci (amplification rate)', y = '% of Samples', color = 'Min Coverage')
```

```{r, fig.width=7, fig.height=5, fig.cap = "**Figure 1:** Sample success rate based on different thresholds of minimum read depth coverage per allele (ASV or amplicon sequence variant observed in the sample). x-axis represents the percentage of loci (amplicon or marker) which contains alleles that pass the threshold. y-axis represents the percentage of samples which amplify a specific percentage of loci given the allele detection threshold mentioned above. The plot represents all analyzed samples (no subsetting based on run and additional variables)."}
plot_precentage_of_samples_over_min_abd
```


Let's do it by Country

```{r}
ReadDepth_coverage = get_ReadDepth_coverage(ampseq_object_abd1, variable = 'Multiplex.cat')

sample_performance = ReadDepth_coverage$plot_read_depth_heatmap$data %>%
      filter(!is.na(var))%>%
      mutate(Read_depth = case_when(
        is.na(Read_depth) ~ 0,
        !is.na(Read_depth) ~ Read_depth
      )) %>%
      summarise(amplified_amplicons1 = sum(Read_depth >= 1)/nrow(ampseq_object_abd1@markers),
                amplified_amplicons5 = sum(Read_depth >= 5)/nrow(ampseq_object_abd1@markers),
                amplified_amplicons10 = sum(Read_depth >= 10)/nrow(ampseq_object_abd1@markers),
                amplified_amplicons20 = sum(Read_depth >= 20)/nrow(ampseq_object_abd1@markers),
                amplified_amplicons50 = sum(Read_depth >= 50)/nrow(ampseq_object_abd1@markers),
                amplified_amplicons100 = sum(Read_depth >= 100)/nrow(ampseq_object_abd1@markers),
                Multiplex.cat = unique(var),
                .by = Sample_id) %>%
      pivot_longer(cols = starts_with('amplified_amplicons'), values_to = 'AmpRate', names_to = 'Threshold') %>%
      mutate(Threshold = as.integer(gsub('amplified_amplicons', '', Threshold)))

plot_precentage_of_samples_over_min_abd_byMultiplex.cat = sample_performance %>%
      summarise(AmpRate5 = round(100*sum(AmpRate >= .05)/n(), 1),
                AmpRate10 = round(100*sum(AmpRate >= .10)/n(), 1),
                AmpRate15 = round(100*sum(AmpRate >= .15)/n(), 1),
                AmpRate20 = round(100*sum(AmpRate >= .20)/n(), 1),
                AmpRate25 = round(100*sum(AmpRate >= .25)/n(), 1),
                AmpRate30 = round(100*sum(AmpRate >= .30)/n(), 1),
                AmpRate35 = round(100*sum(AmpRate >= .35)/n(), 1),
                AmpRate40 = round(100*sum(AmpRate >= .40)/n(), 1),
                AmpRate45 = round(100*sum(AmpRate >= .45)/n(), 1),
                AmpRate50 = round(100*sum(AmpRate >= .50)/n(), 1),
                AmpRate55 = round(100*sum(AmpRate >= .55)/n(), 1),
                AmpRate60 = round(100*sum(AmpRate >= .60)/n(), 1),
                AmpRate65 = round(100*sum(AmpRate >= .65)/n(), 1),
                AmpRate70 = round(100*sum(AmpRate >= .70)/n(), 1),
                AmpRate75 = round(100*sum(AmpRate >= .75)/n(), 1),
                AmpRate80 = round(100*sum(AmpRate >= .80)/n(), 1),
                AmpRate85 = round(100*sum(AmpRate >= .85)/n(), 1),
                AmpRate90 = round(100*sum(AmpRate >= .90)/n(), 1),
                AmpRate95 = round(100*sum(AmpRate >= .95)/n(), 1),
                AmpRate100 = round(100*sum(AmpRate >= 1)/n(), 1),
                .by = c(Threshold, Multiplex.cat)
      ) %>%
      pivot_longer(cols = paste0('AmpRate', seq(5, 100, 5)),
                   values_to = 'Percentage',
                   names_to = 'AmpRate') %>%
      mutate(AmpRate = as.numeric(gsub('AmpRate','', AmpRate)))%>%
      ggplot(aes(x = AmpRate, y = Percentage, color = as.factor(Threshold), group = as.factor(Threshold))) +
      geom_line() +
      geom_vline(xintercept = 50, linetype = 2) +
      facet_wrap(Multiplex.cat~., ncol = 3)+
      theme_bw() +
      labs(x = '% of amplified loci (amplification rate)', y = '% of Samples', color = 'Min Coverage')

```

```{r, include = FALSE}
fig3.height = 3*ceiling(length(unique(plot_precentage_of_samples_over_min_abd_byMultiplex.cat$data$Multiplex.cat))/3)
```

```{r, fig.width=10, fig.height=fig3.height, fig.cap="**Figure 3:** Sample success rate by Sequencing Run based on different thresholds of minimum read depth coverage per allele (ASV or amplicon sequence variant observed in the sample). x-axis represents the percentage of loci (amplicon or marker) which contains alleles that pass the threshold. y-axis represents the percentage of samples which amplify a specific percentage of loci given the allele detection threshold mentioned above."}
plot_precentage_of_samples_over_min_abd_byMultiplex.cat
```

## Identifying likely genotyping errors

After the denoising process of the fastq files, the cigar table might
contains some artefacts that affect downstream analysis. These artifacts
include systematic PCR errors (off-target products and stutters in
INDELs) as well as stochastic PCR errors. All these artifacts coexist
with the true allele within the sample, causing the observation of two
or more alleles (the true allele plus the artifacts) in the particular
loci of the sample where the error has occurred, the increment of the
heterozygousity of the loci (if the error is systematic), and the
increment of polyclonal infections in the population.

However, when analyzing a population we do not expect all samples to be
polyclonal, and it should not be possible to find a locus that is
heterozygous for all samples. Moreover, loci with alternative alleles
only present in heterozygous samples should be rare as they are the
result of *"de-novo"* mutations, and the frequency of heterozygous
samples for that loci should be low because samples comes from different
regions and we do not expect a wide geographic distribution of that
genotype (This might not be true if only one population is been
analyzed).

Here the identification and removal of likely genotyping errors is done
in three sequential steps: 1) Removal of off-target products, 2) masking
of INDELs present in flanking regions of the ASV, 3) and removal of
stochastic PCR errors (alleles mainly present as minor alleles).

The identification of all these likely genotyping errors is based on the
following parameters:

1.  dVSITES_ij: Density of variant sites (SNPs and INDELs) in the allele
    (ASV) $j$ of the locus $i$.

2.  P_ij ($P_{i,j}$): The total number of samples that amplified each
    alternative allele $j$ in locus $i$.

3.  H_ij ($H_{i,j}$): Number of heterozygous samples in the locus $i$
    that carry the alternative allele $j$.

4.  H_ijminor ($H_{i,jminor}$): Number of heterozygous samples in the
    locus $i$ where the alternative allele $j$ is the minor Allele (the
    allele with the lower read counts).

5.  h_ij ($h_{i,j}=H_{i,j}/P_{i,j}$): ratio of heterozygous samples in
    the locus that carry the alternative allele of interest respect to
    the total number of samples that amplified alternative allele.

6.  h_ijminor ($h_{i,jminor}=H_{i,jminor}/H_{i,j}$): ratio of
    heterozygous samples where the alternative allele is the minor
    Allele respect to the total number of heterozygous samples in the
    locus that carry the alternative allele.

7.  p_ij ($p_{i,j}$): population prevalence of the alternative allele
    $j$ in locus $i$.
    
8. flanking_INDEL: Boolean that specify if there are INDELs in the flanking region of the ASV.

9. SNV_in_homopolymer: Boolean that specify if there are SNVs in an homopolymer region. By default and homopolymer is defined as a region of a single nucleotide repeated 5 times in tandem.

10. INDEL_in_homopolymer: Boolean that specify if there are SNVs in an homopolymer region. By default and homopolymer is defined as a region of a single nucleotide repeated 5 times in tandem.

One of the consequences of the geontyping erros, is the increment of heterozygousity in the loci where those errors occurs. Lets check the distribution of the observed heterozygosity by locus, and the distribution of other metrics of the alternative alleles.

```{r}

ref_fasta = '/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/reference/Pviv_P01/PvGTSeq8_refseqs.fasta'

Alt_ASVs_stats = 
  get_ASVs_attributes(ampseq_object, 
                      ref_fasta = ref_fasta)

h_ij_thres = 0.5
h_ijminor_thres = 0.5
Hetob_j_thres = 0.05


plot_Hetob_j = Alt_ASVs_stats %>%
  summarise(Hetob_j = mean(Hetob_j), .by = MHap) %>%
  ggplot(aes(x= Hetob_j)) + 
    geom_vline(xintercept = Hetob_j_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

plot_h_ij = Alt_ASVs_stats %>%
  ggplot(aes(x= h_ij)) + 
    geom_vline(xintercept = h_ij_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

plot_h_ijminor = Alt_ASVs_stats %>%
  ggplot(aes(x= h_ijminor)) + 
    geom_vline(xintercept = h_ijminor_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

Alt_ASVs_metrics_distribution = Alt_ASVs_stats %>%
  mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)),
        
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres)
      )
        
        )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = Hetob_j))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(Hetob_j_cat ~ h_ijminor_cat)+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')

Alt_ASVs_metrics_distribution = ggdraw()+
  draw_plot(plot_Hetob_j, 
                x = 0, width = .4,
                y = 0.66, height = .34)+
  draw_plot(plot_h_ij, 
                x = 0, width = .4,
                y = 0.33, height = .33)+
  draw_plot(plot_h_ijminor, 
                x = 0, width = .4,
                y = 0, height = .33)+
  draw_plot(Alt_ASVs_metrics_distribution, 
                x = .4, width = .6,
                y = 0, height = 1)
      
Alt_ASVs_metrics_distribution
```

### Identification and removal of off-target products

A non-specific product is the result of the amplification of a DNA
template that shares a certain degree of identity in the region of the
primers with the desired target. However, the degree of identity of
these off-targets with respect to the desired target tends to be low,
generating a high density of polymorphisms in the alignment between the
off-target ASV and the reference template. For that reason in this step
we use the density of of variant sites (dVSITES_ij) in order to identify
alleles that are potentially off-target products. The other parameters
explained above can also be included to constraint more the definition.
The filtering criteria can be specified with the argument "mask_formula"
(`maskt_formula = "dVSITES_ij >= 0.3"`).

Run the function `get_ASVs_attributes`:

```{r}

off_target_stats = get_ASVs_attributes(ampseq_object, 
                                       ref_fasta = ref_fasta)
```

Generate a histogram of the density of variant sites per allele.

```{r}

dVSITES_ij_thres = 0.25

plot_off_target_stats = off_target_stats %>%
    ggplot(aes(x= dVSITES_ij)) + 
    geom_vline(xintercept = dVSITES_ij_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

plot_off_target_stats2 = off_target_stats %>%
      mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)
      ),
      dVSITES_ij_cat = case_when(
        dVSITES_ij >= dVSITES_ij_thres ~ 'dVSITES_ij >= 0.25',
        dVSITES_ij < dVSITES_ij_thres ~ 'dVSITES_ij < 0.25'
      ),
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres)
      )
      )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = dVSITES_ij))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(dVSITES_ij_cat ~ Hetob_j_cat + h_ijminor_cat)+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')
    
    plot_off_target_stats = ggdraw()+
      draw_plot(plot_off_target_stats, 
                x = 0, width = .4,
                y = 0, height = 1)+
      draw_plot(plot_off_target_stats2, 
                x = .4, width = .6,
                y = 0, height = 1)

```

```{r, fig.width=8, fig.height=4, fig.cap="**Figure 4:** Identifcation of off-target products. Histogram represents the distribution of the densitity of variant sites in each allele in each locus (dVSITES_ij). The scatter plot shows the distribution of the prevalence of the alternative allele (p_ij), the frequency of the alternative allele as heterozygous (h_ij), the frequency of the alternative allele as minor allele (h_ijminor), and the density of variant sites (dVSITES_ij) of the alternative allele respect to the reference template."}
plot_off_target_stats
```

Identify how many alleles would be removed using a filter of 0.25.

```{r}

n_off_target_alleles = off_target_stats[off_target_stats$dVSITES_ij >= dVSITES_ij_thres,][['Allele']]

print(paste0(length(n_off_target_alleles), ' allele(s) match(es) the criteria to define off-target products'))
```

Check what are those allele:

```{r, eval = FALSE}

View(off_target_stats%>%
  filter(dVSITES_ij >= dVSITES_ij_thres))

```

remove off-target products:

```{r}
ampseq_object = mask_alt_alleles(ampseq_object, mask_formula = 'dVSITES_ij >= 0.25', ref_fasta = ref_fasta)
```

### Products with flanking INDELs

Flanking INDELs are defined as INDELs that occurs at the beginning or
the end of the ASV attached to the primer area. Flanking INDELs can lead
to primer miss binding during PCR by creating additional alleles which
differs in the length of the INDEL in the same sample. The filtering
criteria can be specified in Terra with the argument
"flanking_INDEL_formula" (by default
`"flanking_INDEL_formula": "flanking_INDEL==TRUE&h_ij>=0.66"`).

Identified Flanking INDELs are masked, which means the the internal
region of the ASV is kept to define the allele.

Run the function again `get_ASVs_attributes`:

```{r}
flanking_INDEL_stats = get_ASVs_attributes(ampseq_object, 
                                            ref_fasta = ref_fasta)
```

Generate a scatter plot of the density of variant sites per allele.

```{r}

plot_flanking_INDEL_stats = flanking_INDEL_stats %>%
      mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)),
        flanking_INDEL_cat = case_when(
        flanking_INDEL == TRUE ~ 'w/. flanking INDELs',
        flanking_INDEL == FALSE ~ 'w/o flanking INDELs'
      ),
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres)
      )
        
        )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = Hetob_j))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(flanking_INDEL_cat ~ Hetob_j_cat + h_ijminor_cat)+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')

```

```{r, fig.width=7, fig.height=7, fig.cap="**Figure 5:** Identification of flanking INDELs. The scatter plot shows the distribution of the prevalence of the alternative allele (p_ij), the frequency of the alternative allele as heterozygous (h_ij), and the frequency of the alternative allele as minor allele (h_ijminor). The panels FALSE and TRUE represent alleles that do not contain or contain flanking INDELs."}
plot_flanking_INDEL_stats
```

Identify how many alleles would be removed using
`flanking_INDEL == TRUE`:

```{r}
n_flanking_INDEL_alleles = flanking_INDEL_stats[flanking_INDEL_stats$flanking_INDEL == TRUE, ][['Allele']]
print(paste0(length(n_flanking_INDEL_alleles), ' allele(s) match(es) the criteria to identify products with flanking INDELs'))
```

```{r, eval = FALSE}
View(
  flanking_INDEL_stats %>%
    filter(flanking_INDEL == TRUE)
)

```

Mask alleles with flanking indels:

```{r}

while(length(n_flanking_INDEL_alleles) > 0){
  ampseq_object = mask_alt_alleles(ampseq_object, mask_formula = 'flanking_INDEL == TRUE',
                                      ref_fasta = ref_fasta)
  
  flanking_INDEL_stats2 = get_ASVs_attributes(ampseq_object, 
                                            ref_fasta = ref_fasta)
  n_flanking_INDEL_alleles = 
    flanking_INDEL_stats2[flanking_INDEL_stats2$flanking_INDEL == TRUE, ][['Allele']]
}


```

### PCR errors due to homopolymers

Run the function again `get_ASVs_attributes`:

```{r}
homopolymers_stats = get_ASVs_attributes(ampseq_object, 
                                            ref_fasta = ref_fasta)
```

Generate a scatter plot of the density of variant sites per allele.

```{r}

plot_homopolymers_stats = homopolymers_stats %>%
      mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)
      ),
      
      vsites_in_homopopylmers = case_when(
        SNV_in_homopolymer == FALSE & INDEL_in_homopolymer == FALSE ~ 'No VSites in Hom.',
        SNV_in_homopolymer == TRUE ~ 'SNV in Hom.',
        INDEL_in_homopolymer == TRUE ~ 'INDEL in Hom.'
      ),
      
      vsites_in_homopopylmers = factor(vsites_in_homopopylmers,
                                       levels = c('No VSites in Hom.',
                                                  'SNV in Hom.',
                                                  'INDEL in Hom.')),
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres)
      )
      
      )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = Hetob_j))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(vsites_in_homopopylmers ~ Hetob_j_cat + h_ijminor_cat )+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')

```

```{r, fig.width=7, fig.height=7, fig.cap="**Figure 5:** Identification of flanking INDELs. The scatter plot shows the distribution of the prevalence of the alternative allele (p_ij), the frequency of the alternative allele as heterozygous (h_ij), and the frequency of the alternative allele as minor allele (h_ijminor). The panels FALSE and TRUE represent alleles that do not contain or contain flanking INDELs."}
plot_homopolymers_stats
```

Identify how many alleles would be removed using
`SNV_in_homopolymer == TRUE`:

```{r}
n_snv_in_homopolymer_alleles = homopolymers_stats[
  homopolymers_stats$SNV_in_homopolymer == TRUE & 
    (homopolymers_stats$h_ij > h_ij_thres & homopolymers_stats$h_ijminor > h_ijminor_thres ), ][['Allele']]
print(paste0(length(n_snv_in_homopolymer_alleles), ' allele(s) match(es) the criteria to identify products with SNVs in homopolymers'))
```

Identify how many alleles would be removed using
`INDEL_in_homopolymer == TRUE`:

```{r}
n_indel_in_homopolymer_alleles = homopolymers_stats[homopolymers_stats$INDEL_in_homopolymer == TRUE 
                                                      , ][['Allele']]
print(paste0(length(n_indel_in_homopolymer_alleles), ' allele(s) match(es) the criteria to identify products with INDELs in homopolymers'))
```

Mask alleles with SNVs or INDELs in homopolymers

```{r}
ampseq_object = mask_alt_alleles(ampseq_object, mask_formula = 'INDEL_in_homopolymer == TRUE', ref_fasta = ref_fasta)
```

### Stochastic PCR errors

PCR errors introduce new alleles in both polymorphic and monomorphic
sites. These errors depend on the error rate of the polymerases used in
the sWGA and library generation steps, and they will generate
alternative alleles that are mainly present as the minor allele at a
heterozygous site and their frequency tends to be low in the population.
By analyzing the distribution of population frequency of each
alternative allele p_ij ($p_{i,j}$), the ratio of heterozygous samples
in the locus that carry the alternative allele of interest respect to
the total number of samples that amplified alternative allele h_ij
($h_{i,j}$), and the ratio of heterozygous samples where the alternative
allele is the minor allele respect to the total number of heterozygous
samples h_ijminor ($h_{i,jminor}$), in this step we will remove all
alternative alleles that match our filtering criteria. This filtering
criteria can be specified in Terra with the argument
"PCR_errors_formula" (by default "PCR_errors_formula":
`"h_ij>=0.66&h_ijminor>= 0.66"`)

Use again the function `get_ASVs_attributes`

```{r}
PCR_errors_stats = get_ASVs_attributes(ampseq_object, ref_fasta = ref_fasta)
```

Use an scatter plot to visualize alleles about the thresholds.

```{r}

plot_PCR_errors_stats = PCR_errors_stats %>%
      mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)
      ),
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres))
      )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = Hetob_j))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(Hetob_j_cat~h_ijminor_cat)+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')
```

```{r, fig.width=7, fig.height=5, fig.cap="**Figure 6:** Identification and removal of stochastic PCR errors. The scatter plot shows the distribution of the prevalence of the alternative allele (p_ij), the frequency of the alternative allele as heterozygous (h_ij), and the frequency of the alternative allele as minor allele (h_ijminor). The horizontal panels are defined based on the threshold stablished for h_ijminor."}
plot_PCR_errors_stats
```

How many alleles didn't pass the threshold

```{r}
n_PCR_errors_alleles = PCR_errors_stats[PCR_errors_stats$h_ij >= 0.66 & PCR_errors_stats$h_ijminor >= 0.66, ][['Allele']]
print(paste0(length(n_PCR_errors_alleles), ' allele(s) match(es) the criteria to identify PCR_errors'))
```

```{r, eval = FALSE}
View(
  PCR_errors_stats %>%
    filter(h_ij >= 0.66 & h_ijminor >= 0.66))
                 
```

Remove stochastic PCR errors:

```{r}
ampseq_object = mask_alt_alleles(ampseq_object, mask_formula = 'h_ij >= 0.66 & h_ijminor >= 0.66', ref_fasta = ref_fasta)
```

After removing and masking suspicious alleles, the distribution of the different metrics looks as follows:

```{r}
Alt_ASVs_stats = 
  get_ASVs_attributes(ampseq_object, 
                      ref_fasta = ref_fasta)

plot_Hetob_j = Alt_ASVs_stats %>%
  ggplot(aes(x= Hetob_j)) + 
    geom_vline(xintercept = Hetob_j_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

plot_h_ij = Alt_ASVs_stats %>%
  ggplot(aes(x= h_ij)) + 
    geom_vline(xintercept = h_ij_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

plot_h_ijminor = Alt_ASVs_stats %>%
  ggplot(aes(x= h_ijminor)) + 
    geom_vline(xintercept = h_ijminor_thres,
               linetype = 2) +
    geom_histogram(binwidth = 0.01) + 
    theme_bw()

Alt_ASVs_metrics_distribution = Alt_ASVs_stats %>%
  mutate(h_ijminor_cat = case_when(
        h_ijminor < h_ijminor_thres ~ paste0('h_ijminor < ',h_ijminor_thres),
        h_ijminor >= h_ijminor_thres ~ paste0('h_ijminor >= ',h_ijminor_thres)),
        
      Hetob_j_cat = case_when(
        Hetob_j < Hetob_j_thres ~ paste0('Hetob_j < ',Hetob_j_thres),
        Hetob_j >= Hetob_j_thres ~ paste0('Hetob_j >= ',Hetob_j_thres)
      )
        
        )%>%
      ggplot(aes(x = p_ij, 
                 y = h_ij,
                 color = h_ijminor,
                 size = Hetob_j))+
      geom_point()+
      geom_hline(yintercept = h_ij_thres,
                 linetype = 2) +
      theme_bw()+
      scale_color_continuous(type = 'viridis')+
      facet_grid(Hetob_j_cat~h_ijminor_cat)+
      labs(x = 'Alternative allele prev. (p_ij)',
           y = 'h_ij (H_ij/P_ij)',
           color = 'h_ijminor')

Alt_ASVs_metrics_distribution = ggdraw()+
  draw_plot(plot_Hetob_j, 
                x = 0, width = .4,
                y = 0.66, height = .34)+
  draw_plot(plot_h_ij, 
                x = 0, width = .4,
                y = 0.33, height = .33)+
  draw_plot(plot_h_ijminor, 
                x = 0, width = .4,
                y = 0, height = .33)+
  draw_plot(Alt_ASVs_metrics_distribution, 
                x = .4, width = .6,
                y = 0, height = 1)
      
Alt_ASVs_metrics_distribution
```

## Read depth yield per sample and locus

Inspect again the read depth coverage of our data:

```{r}
ampseq_object@metadata %<>% mutate(Samp_Multiplex = paste(BiologicalRep, Multiplex.cat, sep = '_'))

ReadDepth_coverage = get_ReadDepth_coverage(ampseq_object, variable = 'Samp_Multiplex')
```

```{r, include = FALSE}
Fig5.height = length(unique(ReadDepth_coverage$plot_read_depth_heatmap$data$var))*3
```

#### total read depth heat map
```{r, fig.width=12, fig.height=15, fig.cap="**Figure 7:** Read Coverage per sample per locus. The heatmaps show the number of read-pairs obtained per sample (rows) at each locus (columns). Separate panels are used for each value of Variable1 (e.g., sample geographic origin). Darker reds indicate higher read-depth. Grey indicates lack of signal for the locus/sample."}
ReadDepth_coverage$plot_read_depth_heatmap




heatmap.dat <- ReadDepth_coverage$plot_read_depth_heatmap$data %>% 
  mutate(BiologicalRep = gsub('_.+$', '', var), 
         Multiplex.cat = gsub('^.+_(.+)', '\\1', var),
         Sample_id2 = gsub('-8p', '', paste(str_extract(Sample_id, '^.+\\d+p'), str_extract(Sample_id, '-\\d+$'), sep = '')), 
         Parasitemia = as.numeric(gsub('-', '.', gsub('(^\\w+-|p-.+$)', '', Sample_id2))),
         Sample_id3 = paste(gsub('-.+$', '', Sample_id2), Parasitemia, gsub('^.+-', '', Sample_id2))) %>% arrange(BiologicalRep, Parasitemia) 



heatmap.dat %>% ggplot(aes(x=Markers, y=Sample_id3, fill=Read_depth)) + 
  geom_tile() +
  facet_wrap(~Multiplex.cat, ncol = 3, scales = 'free') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Markers",
    y = "Sample_parasitemia_replicate") +
  scale_fill_gradient(low = "blue1", high = "red") 
 

```

##### assess total read depth performance between multiplexes
```{r}
heatmap.dat %<>% mutate(TechRep = as.integer(gsub('^.+-', '', Sample_id)),
                        Read_depth = case_when(is.na(Read_depth) ~ 0,
                                               .default = Read_depth))

heatmap.dat$Multiplex.cat <- as.factor(heatmap.dat$Multiplex.cat)
bw.dat <- dplyr::filter(heatmap.dat, Multiplex.cat != "NA")

bw.dat$Read_depth_log <- log10(bw.dat$Read_depth + 1)


DBP_1 <- filter(bw.dat, Multiplex.cat != "6plex1" & Multiplex.cat != "4plex1"
                & Markers == "DBP_1")
RBP2b_3 <- filter(bw.dat, Multiplex.cat != "6plex1" & Multiplex.cat != "6plex2"
                & Multiplex.cat != "4plex1" & Markers == "RBP2b_3")
RIPR <- filter(bw.dat, Multiplex.cat != "6plex3" & Markers == "RIPR")
SLARP <- filter(bw.dat, Multiplex.cat != "4plex1" & Multiplex.cat != "4plex2" & Markers == "SLARP")
PIGM <- filter(bw.dat, Markers == "PIGM")
VPS11 <- filter(bw.dat, Multiplex.cat != "6plex3" & Multiplex.cat != "4plex2" &
                  Markers == "VPS11")
FRM1_1 <- filter(bw.dat, Multiplex.cat != "6plex2" & Multiplex.cat != "4plex1" &
                   Multiplex.cat != "4plex2" & Markers == "FRM1_1")
CG2_related <- filter(bw.dat, Multiplex.cat != "4plex2" & Markers == "CG2_related")

cleaned.dat <- rbind(DBP_1, RBP2b_3, RIPR, SLARP, PIGM, VPS11, FRM1_1, CG2_related)
my_comparisons <- list(c("8plex", "6plex1"), c("8plex", "6plex2"), c("8plex", "6plex3"), 
                       c("8plex", "4plex1"), c("8plex", "4plex2"), c("6plex1", "6plex2"),
                       c("6plex1", "6plex3"), c("6plex1", "4plex1"), c("6plex1", "4plex2"),
                       c("6plex2", "6plex3"), c("6plex2", "4plex1"), c("6plex2", "4plex2"),
                       c("6plex3", "4plex1"), c("6plex3", "4plex2"), c("4plex1", "4plex2"))

cleaned.dat %>% ggpaired(x = 'Multiplex.cat', y = 'Read_depth_log', color = 'Multiplex.cat',
      line.color = 'gray', line.size = 0.2, linetype = 2, palette = "jco")+
 stat_compare_means(paired = TRUE, alpha = 0.5)+
 stat_compare_means(comparisons = my_comparisons) +
 scale_y_continuous(breaks = log10(c(0,10^seq(0, 5, 1)) + 1), labels = c(0,10^seq(0, 5, 1)))+
 theme_bw()+
 labs(title = "Total Sample Read depth",
    x = "Assay",
    y = "Read Depth") +
 theme(legend.position = "none",
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(angle = 90, vjust = .5, size = 11),
    axis.title = element_text(size = 13))


```


#### total read depth performance between multiplexes by parasitemia
```{r, height = 12, width = 4}
cleaned.dat <- rbind(DBP_1, RBP2b_3, RIPR, SLARP, PIGM, VPS11, FRM1_1, CG2_related)
custom_legend_title <- "Assay"

cleaned.dat %>% ggpaired(x = 'Multiplex.cat', y = 'Read_depth_log', color = 'Multiplex.cat',facet.by = 'Parasitemia',
      line.color = 'gray', line.size = 0.2, linetype = 2)+
 scale_y_continuous(breaks = log10(c(0,10^seq(0, 5, 1)) + 1), labels = c(0,10^seq(0, 5, 1)))+
 theme_bw()+
 labs(x = "Parasites/μL",
    y = "Read Depth") +
 theme(legend.position = "right",
    axis.text.y = element_text(size = 11),
    axis.text.x = element_blank(),
    axis.ticks.x.bottom = element_blank(),
    axis.title = element_text(size = 13)) +
  scale_color_manual(name = custom_legend_title, values = 
                       scales::hue_pal()(length(unique(cleaned.dat$Multiplex.cat))))
```

#### read depth by marker by parasitemia level by sample
```{r, fig.height= 24, width= 12}
cleaned.dat %>% mutate(Parasitemia2 = as.factor(Parasitemia)) %>% 
  ggplot(aes(x= Multiplex.cat, y = Read_depth_log, color = Multiplex.cat)) +
  geom_boxplot() +
  geom_jitter(width=0.1) + 
  scale_y_continuous(breaks = log10(c(0,10^seq(0, 5, 1)) + 1), labels = c(0,10^seq(0, 5, 1))) +
  labs(x = "Assay",
    y = "Read Depth") +
  facet_wrap(Markers~BiologicalRep+Parasitemia2, ncol = 10) +
  theme(axis.text.x = element_text(angle = 70, size = 20, vjust = 0.5),
        axis.text.y = element_text(size = 20),
        strip.text = element_blank(),
        axis.title = element_text(size = 28),
        legend.position = "none")
```

```{r}
s1_list <- ampseq_object@metadata %>% filter(BiologicalRep == "G4G2463") %>% select(Sample_id) %>% unlist
s2_list <- ampseq_object@metadata %>% filter(BiologicalRep == "HRP0378")  %>% select(Sample_id) %>% unlist

x <- draw_haplotypes(ampseq_object = ampseq_object, comparison_list = list(s1_list, s2_list))

x
```

#### total read depth by marker
```{r}
cleaned.dat <- rbind(DBP_1, RBP2b_3, RIPR, SLARP, PIGM, VPS11, FRM1_1, CG2_related)


cleaned.dat %>% ggpaired(x = 'Multiplex.cat', y = 'Read_depth_log', color = 'Multiplex.cat',
      line.color = 'gray', line.size = 0.2, linetype = 2, palette = "jco", facet.by = "Markers")+
 stat_compare_means(paired = TRUE, alpha = 0.5) +
 scale_y_continuous(breaks = log10(c(0,10^seq(0, 5, 1)) + 1), labels = c(0,10^seq(0, 5, 1)))+
 theme_bw()+
 labs(title = "Total Sample Read depth",
    x = "Multiplex type",
    y = "Log-transformed Read Depth") +
 theme(legend.position = "none",
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(angle = 90, vjust = .5, size = 11),
    axis.title = element_text(size = 13))
```


### Compare Alt_ASV tables to identify minor alleles; then, test to see if genotyping errors or contamination
```{r}
ampseq_object2 = cigar2ampseq(cigar_object, markers = markers, min_abd = 10, min_ratio = 0.1, remove_controls = F) 

ref_fasta2 = '/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/reference/Pviv_P01/PvGTSeq8_refseqs.fasta'

Alt_ASVs_stats2 = 
  get_ASVs_attributes(ampseq_object2, 
                      ref_fasta = ref_fasta2)

ampseq_object2 = mask_alt_alleles(ampseq_object2, mask_formula = 'dVSITES_ij >= 0.25', ref_fasta = ref_fasta2)
ampseq_object2 = mask_alt_alleles(ampseq_object2, mask_formula = 'SNV_in_homopolymer == TRUE', ref_fasta = ref_fasta2)
ampseq_object2 = mask_alt_alleles(ampseq_object2, mask_formula = 'INDEL_in_homopolymer == TRUE', ref_fasta = ref_fasta2)
ampseq_object2 = mask_alt_alleles(ampseq_object2, mask_formula = 'flanking_INDEL == TRUE', ref_fasta = ref_fasta2)


```

```{r}
ampseq_object3 = cigar2ampseq(cigar_object, markers = markers, min_abd = 10, min_ratio = 0.000001, remove_controls = F) 

ref_fasta = '/Users/chh4419/Library/CloudStorage/OneDrive-HarvardUniversity/Rotations/Neafsey/PvSIMPLESeq/MHap-Analysis/docs/reference/Pviv_P01/PvGTSeq8_refseqs.fasta'

Alt_ASVs_stats3 = 
  get_ASVs_attributes(ampseq_object3, 
                      ref_fasta = ref_fasta)

Alt_ASVs_stats3 = 
  get_ASVs_attributes(ampseq_object3, 
                      ref_fasta = ref_fasta, homopolymer_length = 4)

ampseq_object3 = mask_alt_alleles(ampseq_object3, mask_formula = 'dVSITES_ij >= 0.25', ref_fasta = ref_fasta)
ampseq_object3 = mask_alt_alleles(ampseq_object3, mask_formula = 'SNV_in_homopolymer == TRUE', ref_fasta = ref_fasta)
ampseq_object3 = mask_alt_alleles(ampseq_object3, mask_formula = 'INDEL_in_homopolymer == TRUE', ref_fasta = ref_fasta)
ampseq_object3 = mask_alt_alleles(ampseq_object3, mask_formula = 'flanking_INDEL == TRUE', ref_fasta = ref_fasta)
```
# plot ratio distribution
```{r}
max_alleles <- max(sapply(str_split(ampseq_object3@gt, '_'), function(x){length(x)}))

gt = ampseq_object3@gt
major_allele_table = NULL
temp_minor_allele_table = NULL
ratio = NULL

for(clone in 1:max_alleles){
  if(clone == 1){
    major_allele_table = matrix(as.integer(gsub('^.+:', '', gsub('_.+$', '', gt))),
                                ncol = ncol(gt),
                                nrow = nrow(gt),
                                dimnames = list(rownames(gt),
                                                colnames(gt))
                                );
    gt = matrix(gsub('^(\\d|\\w|\\.)+:\\d+_?', '', gt),
                ncol = ncol(gt),
                                nrow = nrow(gt),
                                dimnames = list(rownames(gt),
                                                colnames(gt)))
}else{
  temp_minor_allele_table = matrix(as.integer(gsub('^.+:', '', gsub('_.+$', '', gt))),ncol = ncol(gt),
                                nrow = nrow(gt),
                                dimnames = list(rownames(gt),
                                                colnames(gt)));
  gt = matrix(gsub('^(\\d|\\w|\\.)+:\\d+_?', '', gt),ncol = ncol(gt),
                                nrow = nrow(gt),
                                dimnames = list(rownames(gt),
                                                colnames(gt)));
  ratio = rbind(ratio, temp_minor_allele_table / major_allele_table) 
}}

histogram(c(ratio))

```

```{r}
iSeq_sWGA_Test_ampseq <- read_ampseq("/Users/chh4419/Desktop/iSeq_sWGA_Test_ampseq", format = "csv")
MiSeq_Run_1_ampseq <- read_ampseq("/Users/chh4419/Desktop/MiSeq_Run_1_ampseq", format = "csv")
MiSeq_Test7COI_ampseq <- read_ampseq("/Users/chh4419/Desktop/MiSeq_Test7COI_ampseq", format = "csv")
PvGTSeq_NovaSeq_L1_ampseq <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L1_ampseq", format = "csv")
PvGTSeq_NovaSeq_L2_ampseq <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L2_ampseq", format = "csv")
PvGTSeq_NovaSeq_L2_ampseq_RD1 <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L2_ampseq_RD1", format = "csv")

PvGTSeq_NovaSeq_L1_ampseq@asv_table %>% filter(Amplicon == "PIGM" & (CIGAR == "95T115A" | CIGAR_masked == "95T115A"))

stored_obj = NULL
list_obj = list(iSeq_sWGA_Test_ampseq = iSeq_sWGA_Test_ampseq, 
                MiSeq_Run_1_ampseq = MiSeq_Run_1_ampseq, 
                MiSeq_Test7COI_ampseq = MiSeq_Test7COI_ampseq, 
                PvGTSeq_NovaSeq_L1_ampseq = PvGTSeq_NovaSeq_L1_ampseq, 
                PvGTSeq_NovaSeq_L2_ampseq = PvGTSeq_NovaSeq_L2_ampseq)

# PIGM
for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "PIGM" & 
                  (CIGAR == "95T115A" | CIGAR_masked == "95T115A"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0(CIGAR, 'in', Amplicon, 'was found in', i))
  }
}

for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "PIGM" & 
                  (CIGAR == "1T95T" | CIGAR_masked == "1T95T"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0(CIGAR, 'in', Amplicon, 'was found in', i))
  }
}

for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "PIGM" & 
                  (CIGAR == "95T115A" | CIGAR_masked == "95T115A"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0(CIGAR, 'in', Amplicon, 'was found in', i))
  }
}

# CG2_related
for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "CG2_related" & 
                  (CIGAR == "31C34A100C139G174G184A195G" | CIGAR_masked == "31C34A100C139G174G184A195G"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0(CIGAR, 'in', Amplicon, 'was found in', i))
  }
}

## present
for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "CG2_related" & 
                  (CIGAR == "31C34A100C139G174T184A" | CIGAR_masked == "31C34A100C139G174T184A"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0('31C34A100C139G174T184A', ' in ', 'CG2_related', ' was found in ', i))
  }
}

# FRM1_1
for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "FRM1_1" & 
                  (CIGAR == "1C" | CIGAR_masked == "1C"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0('1C', ' in ', 'FRM1_1', ' was found in ', i))
  }
}

# RIPR
for (i in names(list_obj)){
  temp_obj = list_obj[[i]]
  if (nrow(temp_obj@asv_table %>% 
           filter(Amplicon == "RIPR" & 
                  (CIGAR == "4G76C85G100A" | CIGAR_masked == "4G76C85G100A"))) > 0){
    print('Something happens')
    stored_obj = c(stored_obj, 
                   paste0('4G76C85G100A', ' in ', 'RIPR', ' was found in ', i))
  }
}


## check frequency of CG2_related contamination
PvGTSeq_NovaSeq_L2_ampseq@asv_table %>% filter(CIGAR == "31C34A100C139G174T184A" | CIGAR_masked == "31C34A100C139G174T184A")

sum(grepl("31C34A100C139G174T184A", PvGTSeq_NovaSeq_L2_ampseq@gt[, "CG2_related"]))

PvGTSeq_NovaSeq_L2_ampseq_RD1@asv_table %>% filter(CIGAR == "31C34A100C139G174T184A" | CIGAR_masked == "31C34A100C139G174T184A")

sum(grepl("31C34A100C139G174T184A", PvGTSeq_NovaSeq_L2_ampseq_RD1@gt[, "CG2_related"]))

## see if error PIGM (n=19) is sample specific --> YES. biological, or genotyping error from sWGA
rownames(ampseq_object3@gt)[grepl("92A95T", ampseq_object3@gt[, "PIGM"])] 
```

```{r}
PvGTSeq_NovaSeq_L1_ampseq <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L1_ampseq", format = "csv")
unique(str_extract(PvGTSeq_NovaSeq_L1_ampseq@metadata$Sample_id, pattern = "^."))

dat <- get_polygenomic(PvGTSeq_NovaSeq_L1_ampseq, poly_formula = "Frac_HetLoci >= 0.06 | Fws <= 0.97", update_popsummary = FALSE)
dat2 <- get_polygenomic(p.filtered.dat, poly_formula = "NHetLoci >= 1", update_popsummary = FALSE)

monoclonal.dat <- filter_samples(PvGTSeq_NovaSeq_L1_ampseq, dat$coi_bySample$Clonality == "Monoclonal")
polyclonal.dat <- filter_samples(PvGTSeq_NovaSeq_L1_ampseq, dat$coi_bySample$Clonality == "Polyclonal")

filtered.dat <- filter_loci(monoclonal.dat, monoclonal.dat@markers$amplicon %in% c("PIGM", "RIPR", "VPS11", "CG2_related"))
p.filtered.dat <- filter_loci(polyclonal.dat, polyclonal.dat@markers$amplicon %in% c("PIGM", "RIPR", "VPS11", "CG2_related"))

filtered.dat@gt = gsub("_.+", "", filtered.dat@gt)

amplified.loci <- rowSums(!is.na(filtered.dat@gt))
histogram(amplified.loci)

all4 <- filter_samples(filtered.dat, amplified.loci == 4)

test1 <- sample(all4@metadata$Sample_id, 2, replace = 2)
genotype1 <- gsub(":\\d+$", "", all4@gt[test1[1], ])
genotype2 <- gsub(":\\d+$", "", all4@gt[test1[2], ])
mark.dif <- sum(genotype1 != genotype2)

mark.dif = NULL

for (i in 1:1000){
  test1 <- sample(all4@metadata$Sample_id, 2, replace = FALSE)
  genotype1 <- gsub(":\\d+$", "", all4@gt[test1[1], ])
  genotype2 <- gsub(":\\d+$", "", all4@gt[test1[2], ])
  mark.dif <- c(mark.dif, sum(genotype1 != genotype2))
}

histogram(mark.dif)

##

y <- dat2$coi_bySample$NHetLoci
histogram(y)







dat3 <- get_polygenomic(polyclonal.dat, poly_formula = "NHetLoci >= 10", update_popsummary = FALSE)
View(dat$pop_summary)


x <- get_pop_diversity(all4, strata = 'Run')
```



## PvGTseq samples identified as polyclonal vs. 4plex1
```{r}
PvGTSeq_NovaSeq_L1_ampseq <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L1_ampseq", format = "csv")
unique(str_extract(PvGTSeq_NovaSeq_L1_ampseq@metadata$Sample_id, pattern = "^."))

set1 = PvGTSeq_NovaSeq_L1_ampseq
set1_polyinfo = get_polygenomic(set1, poly_formula = "Frac_HetLoci >= 0.06 | Fws <= 0.97", update_popsummary = FALSE)

set1@metadata = cbind(set1@metadata, set1_polyinfo$coi_bySample)
set2 = filter_samples(set1, set1@metadata$Clonality == "Polyclonal")

set3 = filter_loci(set2, set2@markers$amplicon %in% c("PIGM", "RIPR", "VSP11", "CG2_related"))
set3_polyinfo = get_polygenomic(set3, poly_formula = "NHetLoci >= 1", update_popsummary = FALSE)

set3_polyinfo_0 = set1_polyinfo$coi_bySample %>% filter(Clonality == "Monoclonal")

set1_polyinfo$coi_bySample[["Set"]] = "249"
set3_polyinfo$coi_bySample[["Set"]] = "4"
set3_polyinfo_0[["Set"]] = "4"

set4 = rbind(set1_polyinfo$coi_bySample, set3_polyinfo$coi_bySample, set3_polyinfo_0)

set4 %<>% mutate(Clonality2 = case_when(
  Set == "249"~ Clonality,
  Set == "4" & Sample_id %in% set3_polyinfo_0$Sample_id ~ Clonality,
  Set == "4" &!(Sample_id %in% set3_polyinfo_0$Sample_id) ~ paste(NHetLoci, Clonality, sep = "_")
))

set4 %>% summarise(n_samples = n(), .by = c(Set, Clonality2)) %>% 
  mutate(Clonality2 = factor(Clonality2, levels = c("Polyclonal", "1_Polyclonal", "2_Polyclonal",
                                                    "3_Polyclonal","0_Monoclonal", "Monoclonal"))) %>%
  ggplot(aes(x = Set, y = n_samples, fill = Clonality2)) +
  geom_col()

set5 = filter_loci(set1, set1@markers$amplicon %in% c("PIGM", "RIPR", "VSP11", "CG2_related"))
set5_polyinfo = get_polygenomic(set5, poly_formula = "NHetLoci >= 1", update_popsummary = FALSE)

set5_polyinfo$coi_bySample[["Set"]] = "4.2"
set5_polyinfo$coi_bySample %<>% mutate(Clonality2 = case_when(
  !(Sample_id %in% set3_polyinfo_0$Sample_id) ~ paste(NHetLoci, Clonality, sep = "_"),
  Sample_id %in% set3_polyinfo_0$Sample_id ~ Clonality))


set6 = rbind(set4, set5_polyinfo$coi_bySample)

set6 %>% summarise(n_samples = n(), .by = c(Set, Clonality2)) %>% 
  filter(Set != "4") %>%
  mutate(Clonality2 = factor(Clonality2, levels = c("Polyclonal", "1_Polyclonal", "2_Polyclonal",
                                                    "3_Polyclonal","0_Monoclonal", "Monoclonal"))) %>%
  ggplot(aes(x = Set, y = n_samples, fill = Clonality2)) +
  geom_col() +
  scale_x_discrete(breaks = c("249", "4.2"), labels = c("PvGTseq249", "4plex1"))

```


## Predicting COI using moire
```{r}
if(!require(moire)){
  install.packages("moire", repos = c("https://eppicenter.r-universe.dev", "https://cloud.r-project.org"))
  
  library(moire)
}

## moire supports loading data from either a long format data.frame using load_long_form_data() or from a wide format data.frame using load_delimited_data().

## adjust data to be in appropriate format
PvGTSeq_NovaSeq_L1_ampseq <- read_ampseq("/Users/chh4419/Desktop/PvGTSeq_NovaSeq_L1_ampseq", format = "csv")
unique(str_extract(PvGTSeq_NovaSeq_L1_ampseq@metadata$Sample_id, pattern = "^."))

df <- ampseq2cigar(PvGTSeq_NovaSeq_L1_ampseq)

df_long <- df@cigar_table %>% pivot_longer(cols = all_of(names(df@cigar_table)[-1]), names_to = "alleles", values_to = "counts")

df_long %<>% mutate(locus = gsub(",.+$", "", alleles),
                    allele2 = gsub("^.+,", "", alleles),
                    sample_id = Sample_id) %>%
  select(sample_id, locus, allele2)

df_long$allele = NA

for (locus in unique(df_long$locus)){
  list_of_alleles = unique(df_long[df_long$locus == locus, ][["allele2"]])
  print(paste(list_of_alleles, collapse = ","))
  allele_codes = 1:length(list_of_alleles)
  for (allele in 1:length(list_of_alleles)){
    df_long[df_long$locus == locus & df_long$allele2 == list_of_alleles[allele],][["allele"]] = allele_codes[allele]
  }
  
  print(which(unique(df_long$locus) == locus))
}


write.csv(df_long, "/Users/chh4419/Desktop/df_long.csv")

df_long$allele2 = NULL
data <- load_long_form_data(df_long)

data2 <- load_long_form_data(namibia_data)

## run MCMC
mcmc_results <- moire::run_mcmc(data, is_missing = data$is_missing)


## Revised code
if(!require(moire)){
  install.packages('moire', repos = c('https://eppicenter.r-universe.dev', 'https://cloud.r-project.org'))
  library(moire)
}else{
  library(moire)
}

dat.subset = sample(PvGTSeq_NovaSeq_L1_ampseq@metadata$Sample_id,
                    100)
dat.subset = filter_samples(PvGTSeq_NovaSeq_L1_ampseq, PvGTSeq_NovaSeq_L1_ampseq@metadata$Sample_id %in% dat.subset)

View(cigar_object2@cigar_table)
cigar_object2 <- ampseq2cigar(dat.subset)

cigar_table_long = cigar_object2@cigar_table %>%
  pivot_longer(cols = names(cigar_object2@cigar_table)[-1],
               names_to = 'Cigar_String',
               values_to = 'ReadDepth')

cigar_table_long %<>% filter(!is.na(ReadDepth))
cigar_table_long %<>% mutate(locus = gsub(',.+', '', Cigar_String),
                             allele = gsub('^.+,', '', Cigar_String))
cigar_table_long %<>% rename('Sample_id' = 'sample_id')

length(unique(namibia_data$sample_id))
moire_input <- load_long_form_data(cigar_table_long)

# With data in appropriate format, run MCMC as follows
mcmc_results <- moire::run_mcmc(moire_input, is_missing = moire_input$is_missing)

nrow(cigar_object2@cigar_table)
length(unique(cigar_table_long$sample_id)) ## expect 100
length(unique(cigar_table_long$locus)) ## ~200

coi_data <- mcmc_results$chains$observed_coi
```


## Retention of loci and samples for subsequent analysis

### Locus performance across different countries

```{r}

all_loci_amplification_rate = locus_amplification_rate(ampseq_object, threshold = 0.65, strata = 'Country', update_loci = F)

```

```{r}
fig8.height = 2*length(unique(all_loci_amplification_rate$data$Strata))
```

```{r, fig.width=7, fig.height=fig8.height, fig.cap="**Figure 10:** Locus amplification rate distribution across categories in Variable1. Each panel represents the cateories of Variable1, and the bottom panel represents the total population and the actual number of loci that are retained. Vertical line represents the locus_ampl_rate threshold set by the user."}

all_loci_amplification_rate

```

Keep loci with more than 50% of amplified samples in at least one of the
countries:

```{r}

ampseq_object = locus_amplification_rate(ampseq_object, threshold = 0.5, strata = 'Country', update_loci = T, based_on_strata = T)

```

### Retention of samples for subsequent analysis

Use the function `sample_amplification_rate`:

```{r}
samples_amplification_rate = sample_amplification_rate(ampseq_object, threshold = 0.8, strata = 'Country', update_samples = F)
```

```{r}
fig9.height = 2*length(unique(samples_amplification_rate$data$Strata))
```

```{r, fig.width=7, fig.height=fig9.height, fig.cap="**Figure 11:** Sample amplification rate distribution across categories in Variable1. Each panel represents the categories of Variable1 and the number within each panel represents the number of samples that are retained in each category."}

samples_amplification_rate

```

Keep samples with at least 50% of amplified loci

```{r}
ampseq_object = sample_amplification_rate(ampseq_object, threshold = 0.5, strata = 'Country', update_samples = T)
```

## Export the filtered ampseq_object

As Excel file

```{r}
write_ampseq(ampseq_object, name = '~/Documents/Github/MHap-Analysis/docs/data/Pviv_example/Pviv_ampseq_filtered.xlsx', format = 'excel')
```

The exported ampseq object can also be uploaded in R to be used for
other pipelines

```{r}
ampseq_object2 = read_ampseq(file = '~/Documents/Github/MHap-Analysis/docs/data/Pviv_example/Pviv_ampseq_filtered.xlsx', 
                   format = 'excel')
```

As csv files

```{r}
write_ampseq(ampseq_object, name = '~/Documents/Github/MHap-Analysis/docs/data/Pviv_example/Pviv_ampseq_filtered', format = 'csv')
```

The exported ampseq object can also be uploaded in R to be used for
other pipelines

```{r}
ampseq_object3 = read_ampseq(file = '~/Documents/Github/MHap-Analysis/docs/data/Pviv_example/Pviv_ampseq_filtered', 
                   format = 'csv')
```
